Contextual Combinatorial Cascading Bandits

### 2015-12-07 contextual combinatorial cascading bandits.tex

At time $t$, we denote $\EE_t[\cdot] = \EE[\cdot | \cH_t]$, where $\cH_t$ is the history a time $t$ before the learning agent
	chooses her action. 
By Equation \eqref{eq:expectation}, we have 
$$
\EE[(\gamma_k \bw_{s,\ba_k^s}) | \cH_{s}] = \theta_*^{\top} (\gamma_k x_{s,\ba_k^s}).
$$
Using the ridge regression of data 
$$
\{(\gamma_k x_{s,\ba_k^s}, \gamma_k \bw_{s,\ba_k^s})\}_{k \in[\bO_s], s\in[t]},
$$
we get an $l^2$-regularized least-squares estimate of $\theta_*$ with regularization parameter $\lambda > 0$:
\begin{equation}
\hat{\theta}_t = (\bX_t^{\top}\bX_t + \lambda I)^{-1} \bX_t^{\top} \bY_t,
\end{equation}
where $\bX_t \in \RR^{(\sum_{s=1}^{t}\bO_s) \times d}$ is the matrix whose rows are $\gamma_k x_{s,\ba_k^s}^{\top}$ and $\bY_t$ is the column vector whose elements are $\gamma_k \bw_s(\ba_k^s)$, $k \in[\bO_s], s\in[t]$. Let
$$
\bV_t = \bX_t^{\top} \bX_t + \lambda I = \lambda I + \sum_{s=1}^{t} \sum_{k=1}^{\bO_s} \gamma_k^2 x_{s,\ba_k^s}x_{s,\ba_k^s}^{\top}.
$$
Then $\bV_t \in \RR^{d \times d}$ is a positive invertible matrix. 
In the following, we will mainly use notations for the disjunctive objective when the two cases are similar.

@@

\begin{theorem}[\cite{abbasi2011improved}]
\label{thm:theta_estimate}
Let 
$$
\beta_{t}(\delta) = \sqrt{\ln\left(\frac{\det(\bV_{t})}{\det(\lambda I)}\right) + 2 \ln\left(\frac{1}{\delta}\right)} + \sqrt{\lambda}.
$$
\wei{I thought the following is simpler. Should we use the following?}
$$
\beta_{t}(\delta) = \sqrt{\ln\left(\frac{\det(\bV_{t})}{\lambda^d \delta^2}\right)} + \sqrt{\lambda}.
$$
Then for any $\delta > 0$, with probability at least $1 - \delta$, for all $t > 0$, we have
\begin{equation}
\label{eq:estimateTheta}
\norm{\hat{\theta}_t - \theta_{\ast}}_{\bV_{t}} \leq \beta_{t}(\delta).
\end{equation}
\end{theorem}

@@

\STATE {//Update statistics}
\STATE {$\bV_{t} \leftarrow \bV_{t-1} + \sum_{k=1}^{\bO_t} \gamma_k^2 x_{t, \ba_k^t}x_{t, \ba_k^t}^{\top}$}
\STATE {$\bX_t \leftarrow [\bX_{t-1}; ~\gamma_1 x_{t, \ba_1^t}^{\top};  ~\ldots; ~\gamma_{\bO_t} x_{t, \ba_{\bO_t}^t}^{\top}]$}
\STATE {$\bY_t \leftarrow [\bY_{t-1}; ~\gamma_1 \bw(\ba_1^t); ~\ldots; ~\gamma_{\bO_t} \bw_t(\ba_{\bO_t}^t)]$}
\STATE {$\hat{\theta}_t \leftarrow (\bX_t^\top \bX_t + \lambda I)^{-1} \bX_{t}^\top \bY_t$}
\STATE {$\beta_{t}(\delta) \leftarrow \sqrt{\ln(\det(\bV_{t})) + 2 \ln(\frac{1}{\delta})} + \sqrt{\lambda}$ \wei{This $\beta$ is different from what is defined in Theorem 4.1.}}
\ENDFOR {~~$t$}
\end{algorithmic}
\end{algorithm}

@@

\subsection{Proofs}

@@

The following lemma provides the relationship between an estimated mean vector and its
	upper bound, in terms of their $f$ function values.
\wei{The symbol here may need to be changed. $w_i$ coincides with weight $w_i$ before, 
		and $p$ coincides with many related probability definition using $p$.
	$w_i$ can be replaced with $u_i$ (standing for upper bound), 
	$u_i$ can be replaced with $r_i$ (standing for radius), but not sure yet
	about $p_i$.}
\shuai{Maybe we can replace $p_i$ by $w_i$ like the lemma above. I have changed accordingly.}
\begin{lemma}
\label{lem:estimateTech}
Suppose $0 \leq \gamma_k, \gamma_k' \leq 1, k \in [K]$. Let $0 \leq w_1, \ldots, w_m \leq 1$, $r_1, \ldots, r_m \geq 0$ and $u_i = \min\{w_i + r_i, 1\}, 1 \leq i \leq m$. Then
\begin{align*}
&f(A^{(m)}, u) \leq f(A^{(m)}, w) + \sum_{k=1}^{m} \gamma_k r_k,\\
&f_{\wedge}(A^{(m)}, u) \leq f_{\wedge}(A^{(m)}, w) + \sum_{k=1}^{m} (1 - \gamma_k') r_k,
\end{align*}
where $A^{(m)} = (1, 2, \ldots, m), u = (u_1, \ldots, u_m)$ and $w = (w_1, \ldots, w_m)$.
\end{lemma}
\begin{proof}
We prove this by induction. It holds obviously for $m = 1$.
\begin{align*}
&~~f(A^{(m+1)}, u)\\
&= f(A^{(m)}, u) + \gamma_{m+1}\prod_{k=1}^m(1 - u_k) u_{m+1}\\
&\leq f(A^{(m)}, u) +  \gamma_{m+1} \prod_{k=1}^m(1 - w_k) (w_{m+1} + r_{m+1})\\
&\leq f(A^{(m)}, w) + \sum_{k=1}^m \gamma_k r_k \\
&\qquad + \gamma_{m+1} \prod_{k=1}^m(1 - w_k) w_{m+1} + \gamma_{m+1} r_{m+1}\\
&= f(A^{(m+1)}, w) + \sum_{k=1}^{m+1} \gamma_k r_k
\end{align*}
\begin{align*}
&~~f_{\wedge}(A^{(m+1)}, u) \\
&= f_{\wedge}(A^{(m)}, u) -\prod_{k=1}^{m} u_k \\
&\qquad+ \gamma_{m+1}' (\prod_{k=1}^{K} u_k) (1 - u_{K+1})+ \prod_{k=1}^{K+1} u_k\\
&= f_{\wedge}(A^{(m)}, u) - (1 - \gamma_{m+1}') (\prod_{k=1}^{m} u_k) (1 - u_{m+1})\\
&\leq f_{\wedge}(A^{(m)}, u) - (1 - \gamma_{m+1}') (\prod_{k=1}^{m} w_k) (1 - u_{m+1})\\
&\leq f_{\wedge}(A^{(m)}, u) -(1 - \gamma_{m+1}') (\prod_{k=1}^{m} w_k)  (1 - w_{m+1} - r_{m+1})\\
&\leq f_{\wedge}(A^{(m)}, w) +  \sum_{k=1}^{m} (1 - \gamma_k') r_k \\
&\qquad - (1 - \gamma_{m+1}') (1 - w_{m+1}) \prod_{k=1}^{m} w_k + (1 - \gamma_{m+1}') r_{m+1}\\
&\leq f_{\wedge}(A^{(m+1)}, w) + \sum_{k=1}^{m+1} (1 - \gamma_k') r_k
\end{align*}
\end{proof}

\begin{proof}[ of Lemma \ref{lem:DeltaEstimate}]
By Lemma \ref{lem:estimateU}, $\theta_{\ast}^{\top}x_t \leq \bU_t$. Then by Lemma \ref{lem:increasing},
$$
f_t^{\ast} = f(A_t^{\ast}, \theta_{\ast}^{\top}x_t) \leq f(A_t^{\ast}, \bU_t).
$$
By Lemma \ref{lem:estimateTech} and the definition of $\bU_t$ \eqref{eq:defU},
$$
f(A, \bU_t) \leq f(A, \theta_{\ast}^{\top}x_t) + \sum_{k=1}^{\abs{A}} \gamma_k \beta_{t-1}(\delta)\norm{x_{t, a_k^t}}_{\bV_{t-1}^{-1}}.
$$
Because $f(A_t^{\ast}, \bU_t) \leq f(A, \bU_t)$, then 
$$
f_t^{\ast} \leq f(A, \theta_{\ast}^{\top}x_t) + \sum_{k=1}^{\abs{A}} \gamma_k \beta_{t-1}(\delta)\norm{x_{t, a_k^t}}_{\bV_{t-1}^{-1}}.
$$
Therefore,
$$
R(t, A) = f_t^{\ast} - f(A, \theta_{\ast}^{\top}x_t) \leq \sum_{k=1}^{\abs{A}} \gamma_k \beta_{t-1}(\delta)\norm{x_{t, a_k^t}}_{\bV_{t-1}^{-1}}.
$$
The result for the disjunctive case can be deduced similarly.
\end{proof}

>>>>

At time $t$, we denote $\EE_t[\cdot] = \EE[\cdot | \cH_t]$, where $\cH_t$ is the history a time $t$ before the learning agent
	chooses her action. 
By Equation \eqref{eq:expectation}, we have 
$$
\EE[(\gamma_k \bw_{s,\ba_k^s}) | \cH_{s}] = \theta_*^{\top} (\gamma_k x_{s,\ba_k^s}).
$$
Using the ridge regression of data 
$$
\{(\gamma_k x_{s,\ba_k^s}, \gamma_k \bw_{s,\ba_k^s})\}_{k \in[\bO_s], s\in[t]},
$$
we get an $l^2$-regularized least-squares estimate of $\theta_*$ with regularization parameter $\lambda > 0$:
\begin{equation}
\hat{\theta}_t = (\bX_t^{\top}\bX_t + \lambda I)^{-1} \bX_t^{\top} \bY_t,
\end{equation}
where $\bX_t \in \RR^{(\sum_{s=1}^{t}\bO_s) \times d}$ is the matrix whose rows are $\gamma_k x_{s,\ba_k^s}^{\top}$ and $\bY_t$ is the column vector whose elements are $\gamma_k \bw_s(\ba_k^s)$, $k \in[\bO_s], s\in[t]$. Let
$$
\bV_t = \bX_t^{\top} \bX_t + \lambda I = \lambda I + \sum_{s=1}^{t} \sum_{k=1}^{\bO_s} \gamma_k^2 x_{s,\ba_k^s}x_{s,\ba_k^s}^{\top}.
$$
Then $\bV_t \in \RR^{d \times d}$ is a symmetric positive definite matrix. 
For any symmetric positive definite matrix $V \in \RR^{d \times d} $ and any
	vector $x \in \RR^{d \times 1}$, we define the $2$-norm of $x$ based on
	$V$ to be $\norm{x}_V = (x^{\top} V x)^{\frac{1}{2}}$.
In the following, we will mainly use notations for the disjunctive objective when the two cases are similar.

@@

\begin{theorem}[\cite{abbasi2011improved}]
\label{thm:theta_estimate}
Let 
%$$
%\beta_{t}(\delta) = \sqrt{\ln\left(\frac{\det(\bV_{t})}{\det(\lambda I)}\right) + 2 \ln\left(\frac{1}{\delta}\right)} + \sqrt{\lambda}.
%$$
%\wei{I thought the following is simpler. Should we use the following?}
$$
\beta_{t}(\delta) = \sqrt{\ln\left(\frac{\det(\bV_{t})}{\lambda^d \delta^2}\right)} + \sqrt{\lambda}.
$$
Then for any $\delta > 0$, with probability at least $1 - \delta$, for all $t > 0$, we have
\begin{equation}
\label{eq:estimateTheta}
\norm{\hat{\theta}_t - \theta_{\ast}}_{\bV_{t}} \leq \beta_{t}(\delta).
\end{equation}
\end{theorem}

@@

\STATE {//Update statistics}
\STATE {$\bV_{t} \leftarrow \bV_{t-1} + \sum_{k=1}^{\bO_t} \gamma_k^2 x_{t, \ba_k^t}x_{t, \ba_k^t}^{\top}$}
\STATE {$\bX_t \leftarrow [\bX_{t-1}; ~\gamma_1 x_{t, \ba_1^t}^{\top};  ~\ldots; ~\gamma_{\bO_t} x_{t, \ba_{\bO_t}^t}^{\top}]$}
\STATE {$\bY_t \leftarrow [\bY_{t-1}; ~\gamma_1 \bw(\ba_1^t); ~\ldots; ~\gamma_{\bO_t} \bw_t(\ba_{\bO_t}^t)]$}
\STATE {$\hat{\theta}_t \leftarrow (\bX_t^\top \bX_t + \lambda I)^{-1} \bX_{t}^\top \bY_t$}
\STATE {$\beta_{t}(\delta) \leftarrow \sqrt{\ln(\det(\bV_{t})/(\lambda^d \delta^2))} + \sqrt{\lambda}$ }
\ENDFOR {~~$t$}
\end{algorithmic}
\end{algorithm}

@@

\subsection{Proof of Theorem~\ref{thm:main}}

@@

The following lemma provides the relationship between an estimated mean vector $w$
	and its upper bound vector $u$, in terms of their $f$ function values.
\begin{lemma}
\label{lem:estimateTech}
Suppose $0 \leq \gamma_k, \gamma_k' \leq 1, k \in [K]$. Let $0 \leq w_1, \ldots, w_m \leq 1$, $r_1, \ldots, r_m \geq 0$ and $u_i = \min\{w_i + r_i, 1\}, 1 \leq i \leq m$. Then
\begin{align*}
&f(A^{(m)}, u) \leq f(A^{(m)}, w) + \sum_{k=1}^{m} \gamma_k r_k,\\
&f_{\wedge}(A^{(m)}, u) \leq f_{\wedge}(A^{(m)}, w) + \sum_{k=1}^{m} (1 - \gamma_k') r_k,
\end{align*}
where $A^{(m)} = (1, 2, \ldots, m), u = (u_1, \ldots, u_m)$ and $w = (w_1, \ldots, w_m)$.
\end{lemma}
\begin{proof}
We prove this by induction. It holds obviously for $m = 1$.
\begin{align*}
&~~f(A^{(m+1)}, u)\\
&= f(A^{(m)}, u) + \gamma_{m+1}\prod_{k=1}^m(1 - u_k) u_{m+1}\\
&\leq f(A^{(m)}, u) +  \gamma_{m+1} \prod_{k=1}^m(1 - w_k) (w_{m+1} + r_{m+1})\\
&\leq f(A^{(m)}, w) + \sum_{k=1}^m \gamma_k r_k \\
&\qquad + \gamma_{m+1} \prod_{k=1}^m(1 - w_k) w_{m+1} + \gamma_{m+1} r_{m+1}\\
&= f(A^{(m+1)}, w) + \sum_{k=1}^{m+1} \gamma_k r_k
\end{align*}
\begin{align*}
&~~f_{\wedge}(A^{(m+1)}, u) \\
&= f_{\wedge}(A^{(m)}, u) -\prod_{k=1}^{m} u_k \\
&\qquad+ \gamma_{m+1}' (\prod_{k=1}^{m} u_k) (1 - u_{m+1})+ \prod_{k=1}^{m+1} u_k\\
&= f_{\wedge}(A^{(m)}, u) - (1 - \gamma_{m+1}') (\prod_{k=1}^{m} u_k) (1 - u_{m+1})\\
&\leq f_{\wedge}(A^{(m)}, u) - (1 - \gamma_{m+1}') (\prod_{k=1}^{m} w_k) (1 - u_{m+1})\\
&\leq f_{\wedge}(A^{(m)}, u) -(1 - \gamma_{m+1}') (\prod_{k=1}^{m} w_k)  (1 - w_{m+1} - r_{m+1})\\
&\leq f_{\wedge}(A^{(m)}, w) +  \sum_{k=1}^{m} (1 - \gamma_k') r_k \\
&\qquad - (1 - \gamma_{m+1}') (1 - w_{m+1}) \prod_{k=1}^{m} w_k + (1 - \gamma_{m+1}') r_{m+1}\\
&\leq f_{\wedge}(A^{(m+1)}, w) + \sum_{k=1}^{m+1} (1 - \gamma_k') r_k
\end{align*}
\end{proof}

The following lemma provides the important result on the regret at each time $t$
	in terms of feature vectors of base arms in the round.
\begin{lemma}
	\label{lem:DeltaEstimate}
	For any time $t$ and $A = (a_1, \ldots, a_{\abs{A}})$, if $f(A_t^*, \bU_t) \leq f(A, \bU_t)$, then we have
	$$
	R(t,A) \leq \sum_{k=1}^{\abs{A}} \gamma_k \beta_{t-1}(\delta)\norm{x_{t,a_k}}_{\bV_{t-1}^{-1}};
	$$
	if $f_{\wedge}(A_t^*, \bU_t) \leq f_{\wedge}(A, \bU_t)$, then we have
	$$
	R_{\wedge}(t, A) \leq \sum_{k=1}^{\abs{A}} \gamma_k \beta_{t-1}(\delta)\norm{x_{t,a_k}}_{\bV_{t-1}^{-1}}.
	$$
\end{lemma}
\begin{proof}
By Lemma \ref{lem:estimateU}, $\theta_{\ast}^{\top}x_t \leq \bU_t$. Then by Lemma \ref{lem:increasing},
$$
f_t^{\ast} = f(A_t^{\ast}, \theta_{\ast}^{\top}x_t) \leq f(A_t^{\ast}, \bU_t).
$$
By Lemma \ref{lem:estimateTech} and the definition of $\bU_t$ \eqref{eq:defU},
$$
f(A, \bU_t) \leq f(A, \theta_{\ast}^{\top}x_t) + \sum_{k=1}^{\abs{A}} \gamma_k \beta_{t-1}(\delta)\norm{x_{t, a_k^t}}_{\bV_{t-1}^{-1}}.
$$
Because $f(A_t^{\ast}, \bU_t) \leq f(A, \bU_t)$, then 
$$
f_t^{\ast} \leq f(A, \theta_{\ast}^{\top}x_t) + \sum_{k=1}^{\abs{A}} \gamma_k \beta_{t-1}(\delta)\norm{x_{t, a_k^t}}_{\bV_{t-1}^{-1}}.
$$
Therefore,
$$
R(t, A) = f_t^{\ast} - f(A, \theta_{\ast}^{\top}x_t) \leq \sum_{k=1}^{\abs{A}} \gamma_k \beta_{t-1}(\delta)\norm{x_{t, a_k^t}}_{\bV_{t-1}^{-1}}.
$$
The result for the disjunctive case can be deduced similarly.
\end{proof}
